A flexible integer linear programming formulation for scheduling clinician on-call service in hospitals

### 2020-07-21 paper/paper.comparison.tex
Add another paper for comparison + minor fixes

Prior work by [authors] have   %avoid editorial type statements without citations or numbers, best to state as objective statements
provided categories to organize the types of problems and approaches
commonly used for real-world nurse rostering.
These categories highlight the similarities
and differences
among computationally easy and hard instances of nurse rostering and 
help end-users determine
which algorithms
are best suited to each type of problem.

In 2010, De Causmaecker and Berghe [ref] addressed a gap in the literature 
surrounding a central categorization system for the nurse rostering problems identified in the literature.
The authors created a categorization framework that takes into account
hard and soft constraints related to nurse preferences, consecutive
assignments, coverage and various flexibilities in terms of shift and 
skill types. The described framework is inspired by the categorization used
in theoretic scheduling problems which focuses on three main aspects of scheduling:
(1) the machine environment;
(2) the job characteristics;
and (3) the objective function.
De Causmaecker and Berghe use these three aspects as a basis for their categorization,
and adjust them to fit the realm of nurse rostering. A description of the categories defined
by the authors is given in Table \ref{tbl:nurse-rostering-categorization}.

@@

%The authors categorized
%nurse rostering problems using a similar approach used in theoretic scheduling problems,
%based around the machine environment, job characteristics and constraints,
%and optimization objective. They also created a classification that takes into account
%hard and soft constraints related to nurse preferences, consecutive 
%assignments, coverage and various flexibilities in terms of shift and
%skill types.

%De Causmaecker and Berghe went on to describe where recent examples of the nurse rostering
%literature fit into their classification. %SM: give examples of their recent examples.

@@

%[maybe 1 additional paragraph on categorization systems?] %SM: yes, describe a bit more or a table/box as the list of categories. That way - flows into next paragraph on where to place the SMH solution.
%SM i.e. describe/define each category

@@

De Causmaecker and Berghe also present examples from the nurse rostering literature
and their appropriate categorization in their framework. We use some of these examples
to compare our clinician
scheduling problem and solution approach with existing literature. Azaiez and Al Sharif [ref]
describe a nurse rostering problem to schedule nurses on a per-day basis, for a total period
of 28 days. Azaiez and Al Sharif's problem is categorized as $ASBN/V2/PL$.
Similar to our clinician scheduling problem, the authors include hard constraints
to prevent consecutive assignments, and enforce a minimum and maximum number of shifts for each nurse.
Unlike our problem, Azaiez and Al Sharif choose an optimization goal that focuses on
balancing the workload among nurses, preventing day-night consecutive shifts and 
avoiding off-on-off assignments. The authors do not attempt to incorporate nurse preferences
for time off into the model.

In [ref-Bilgin 2008], Bilgin et al address a nurse rostering problem using local search.
De Causmaecker and Berghe categorize the problem in [ref-Bilgin] as $ASBI/RVNO/PLR$.
Bilgin et al incorporate coverage constraints, overlapping assignment constraints
and horizontal constraints, which include constraints arising from employment contracts
(such as minimum and maximum thresholds).
Unlike our approach for these constraints, the authors consider coverage constraints to be soft
constraints, in order to prevent over-constraining of the problem.
Bilgin et al also incorporate nurse requests into the model formulation while allowing
nurses to request request days off and days on.

[...]

>>>>

%Prior work by [authors] have   %avoid editorial type statements without citations or numbers, best to state as objective statements
%provided categories to organize the types of problems and approaches
%commonly used for real-world nurse rostering.
%These categories highlight the similarities
%and differences
%among computationally easy and hard instances of nurse rostering and 
%help end-users determine
%which algorithms
%are best suited to each type of problem.

In 2010, De Causmaecker and Berghe~\cite{de_causmaecker_categorisation_2011} addressed a gap in the literature 
surrounding a central categorization system for the nurse rostering problems identified in the literature.
The authors created a categorization framework that takes into account
hard and soft constraints related to nurse preferences, consecutive
assignments, coverage and various flexibilities in terms of shift and 
skill types. The described framework is inspired by the categorization used
in theoretic scheduling problems which focuses on three main aspects of scheduling:
(1) the machine environment;
(2) the job characteristics;
and (3) the objective function.
De Causmaecker and Berghe use these three aspects as a basis for their categorization,
and adjust them to fit the realm of nurse rostering. A description of the categories defined
by the authors is given in Table \ref{tbl:nurse-rostering-categorization}.

@@

@@

@@

De Causmaecker and Berghe also present examples from the nurse rostering literature
and their appropriate categorization in their framework. We use some of these examples
to compare our clinician
scheduling problem and solution approach with existing literature. Azaiez and Al Sharif~\cite{azaiez_0-1_2005}
describe a nurse rostering problem to schedule nurses on a per-day basis, for a total period
of 28 days. Azaiez and Al Sharif's problem is categorized as $ASBN/V2/PL$.
Similar to our clinician scheduling problem, the authors include hard constraints
to prevent consecutive assignments, and enforce a minimum and maximum number of shifts for each nurse.
Unlike our problem, Azaiez and Al Sharif choose an optimization goal that focuses on
balancing the workload among nurses, preventing day-night consecutive shifts and 
avoiding off-on-off assignments. The authors do not attempt to incorporate nurse preferences
for time off into the model.

In~\cite{bilgin_local_2012}, Bilgin et al address a nurse rostering problem using local search.
De Causmaecker and Berghe categorize the problem in~\cite{bilgin_local_2012} as $ASBI/RVNO/PLR$.
Bilgin et al incorporate coverage constraints, overlapping assignment constraints
and horizontal constraints, which include constraints arising from employment contracts
(such as minimum and maximum thresholds).
Unlike our approach for these constraints, the authors consider coverage constraints to be soft
constraints, in order to prevent over-constraining of the problem.
Bilgin et al also incorporate nurse requests into the model formulation by allowing
nurses to request days off and days on.

Burke et al~\cite{burke_metaheuristics_2006} present a flexible framework for nurse rostering based on various approaches
observed from Belgin hospitals. 
In~\cite{de_causmaecker_categorisation_2011}, Burke et al's work is categorized as $ASBCI/RVNTO/PR$.
The authors use a two-step algorithm comprised of an initialization procedure
which provides a solution satisfying only the hard constraints of their problem,
and a metaheuristic procedure, a hybrid tabu search, which aim to satisfy as many soft constraints as possible.
Unlike our clinician scheduling problem, 
Burke et al's hard constraints address only coverage requirements in an effort to make the first
step of their algorithm efficient.
In their metaheurstic procedure, the authors optimize soft constraints including assignment of free weekends
and minimum number of assignments per staff member.
Notably, Burke et al's approach make use of ``time interval personnel requirements'', which describe
the amount of required staff per skill category as a varying number throughout the work day.

### 2020-07-21 paper/paper.intro.tex
Add another paper for comparison + minor fixes

We begin by describing the details of the problem in Section~\ref{sec:problem}, and presenting
our ILP formulation in Section~\ref{sec:methods}. Next, we compare the results
of our formulation to manually-created schedules, and evaluate the performance
of the algorithm on simulated data in Section~\ref{sec:experiments}. Finally, we
discuss and interpret the results in Section~\ref{sec:discussion}.

>>>>

We begin by describing the details of the problem in Section~\ref{sec:problem}, and presenting
our ILP formulation in Section~\ref{sec:methods}. We then categorize our nurse rostering
problem and solution approach in the categorization scheme of~\cite{de_causmaecker_categorisation_2011}
and compare our problem to recent studies in Section~\ref{sec:comparison}.
Next, we analyze the effectiveness of our automated approach when compared
to manually-created schedules, and evaluate the performance
of the algorithm on simulated data in Section~\ref{sec:experiments}. Finally, we
discuss and interpret the results in Section~\ref{sec:discussion}.
